---
title: "Severe weather events impacts on the US communities and municipalities"
author: "Eddy ∆"
date: "30/04/2017"
subtitle: 'Reproducible Research: Peer Assessment 2'
keep_md: yes
---

## Synopsis

This report aim is to provide an analysis of severe weather events impacts on the US communities and municipalities, from both public health and economic problems perspective.
We will study fatalities, injuries, property,  and crop damage weather event can generate.
A better understanding of such impacts can help to prevent or handle more efficiently such outcomes.
To do so, we will explore the U.S. National Oceanic and Atmospheric Administration's (NOAA) storm database.

From those information we'll try to identify :

1. Which types of events (as indicated in the **EVTYPE** variable) are most harmful with respect to population health across the United States.
2. Which types of events have the greatest economic consequences across the United States.

## Data Processing

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load_file}
setwd('~/projects/training/datascience/reproducible_research/storm')

get_file <- function(file_url = 'https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2', data_dir = 'data'){
    if(!dir.exists(data_dir)) dir.create(data_dir)
    file_name <- basename(file_url)
    file_path <- file.path(data_dir, file_name)
    if(!file.exists(file_path))
        download.file(file_url, file_path, quiet = TRUE)
    if(!file.exists(file_path))
        stop("Unable to download file ", file_url)
    # return the archive file name
    file_name
}
# if (!file.exists('load_data.R')) setwd('~/projects/training/datascience/reproducible_research/storm')
# source('load_data.R')

# Load data from files
file_archive <- get_file()
```

```{r storm}
if(exists('storm')){
    tr <- c(H = '2', h = '2', K = '3', M = '6', m = '6', B = '9', '+' = '1', '-' = '-1', '?' = NA, '0' = '0', '1' = '1', '2' = '2', '3' = '3', '4' = '4', '5' = '5', '6' = '6', '7' = '7', '8' = '8')
    exp_value <- function(val, exp) val * 10^as.numeric(tr[exp])
    get_year <- function(date, time, time_zone) as.numeric(format(strptime(paste(gsub('([0-9/]+).+', '\\1', date), time, time_zone), format = '%m/%d/%Y %H%M'), '%Y'))
    storm <- transform(
        read.csv(file.path('data', file_archive), sep = ',', stringsAsFactors = FALSE, na.strings = c('', 'NA')),
        # create begin and end date type and years
        BGN_YEAR = get_year(BGN_DATE, BGN_TIME, TIME_ZONE),
        END_YEAR = get_year(END_DATE, END_TIME, TIME_ZONE),
        # get the cost on the same scale
        PROP = mapply(exp_value, PROPDMG, PROPDMGEXP),
        CROP = mapply(exp_value, CROPDMG, CROPDMGEXP)
        # change the 'COUNTYNAME', 'STATE', 'STATEOFFIC', 'EVTYPE', 'ZONENAMES', 'WFO', 'BGN_LOCATI', and 'END_LOCATI' column type to factor
        # COUNTYNAME = mapply(factor, COUNTYNAME),
        # STATE = mapply(factor, STATE),
        # STATEOFFIC = mapply(factor, STATEOFFIC),
        # EVTYPE = mapply(factor, EVTYPE),
        # ZONENAMES = mapply(factor, ZONENAMES),
        # WFO = mapply(factor, WFO),
        # BGN_LOCATI = mapply(factor, BGN_LOCATI),
        # END_LOCATI = mapply(factor, END_LOCATI)
    )
    cols <- c('COUNTYNAME', 'STATE', 'STATEOFFIC', 'EVTYPE', 'ZONENAMES', 'WFO', 'BGN_LOCATI', 'END_LOCATI')
    storm[, cols] <- lapply(storm[, cols, drop = FALSE], as.factor)
}
```

```{r year_range}
year_range <- range(storm$BGN_YEAR, na.rm = T)
```

The dataset contains `r nrow(storm)` rows and `r length(names(storm))` columns.
And log events from `r year_range[1]` to `r year_range[2]`.

Let's first have a look at the number of events distribution across the years in United States.
```{r event_count}
# Build a boxplot to have an overview of the data distribution
# hist(storm$BGN_YEAR, main = 'Events in United States', xlab = 'Years', ylab = 'Events (nb)')
boxplot(storm$BGN_YEAR, main = 'Events in United States', ylab = 'Years')
evt_count <- aggregate(BGN_DATE ~ BGN_YEAR, data = storm, FUN = length)
evt_count_range <- range(evt_count$BGN_DATE, na.rm = T)
```

We can notice a constant increase of the count of events across the year.
We are able to confirm the lack of recorded event in the early years in the dataset (from `r evt_count_range[1]` to `r evt_count_range[2]`).

### Analysis of Severe weather events impact on population health across the United States

Let's now at the impact of the events on the population in terms of **INJURIES** and **FATALITIES**.
First we need to have an idea of the most harmful events.
To do so we'll have a look at the total numbers of  **Injuries** and **Fatalities** per event types.

```{r harm}
injuries <- aggregate(INJURIES ~ EVTYPE, storm, sum, na.rm = T)
fatalities <- aggregate(FATALITIES ~ EVTYPE, storm, sum, na.rm = T)
# put side by side the costs
harm <- merge(injuries, fatalities, by = 'EVTYPE')
harm$CASES <- harm$INJURIES + harm$FATALITIES
# sort the tables
injuries <- injuries[order(injuries$INJURIES, decreasing = T), ]
fatalities <- fatalities[order(fatalities$FATALITIES, decreasing = T), ]
harm <- harm[order(harm$CASES, decreasing = T), ]
```

```{r harm_plot}
# reshape the table columns in a single column type
top_harm <- reshape(head(harm, 20), direction = 'long', varying = c('INJURIES', 'FATALITIES', 'CASES'), v.names = 'cases', timevar = 'type', times = c('Injuries', 'Fatalities', 'Total'))
library(ggplot2)
ggplot(data = top_harm, aes(x = EVTYPE, y = cases, group = type, colour = type)) +
    geom_line() + geom_point() +
    labs(title = 'Weather events impact on population health', x = 'Event types', y = 'Number of fatalities and injuries') +
    theme(axis.text.x=element_text(angle = 90, hjust = 1))
```

```{r harm_tables}
head(harm, 20)
head(fatalities, 10)
head(injuries, 10)
```

**`r harm$EVTYPE[1]`** are by far the most harmful event, followed by **`r harm$EVTYPE[2]`** and **`r harm$EVTYPE[3]`**.

**`r injuries$EVTYPE[1]`** cost the most injuries, followed by **`r injuries$EVTYPE[2]`** and **`r injuries$EVTYPE[3]`**.

**`r fatalities$EVTYPE[1]`**  the most lethals event, followed by **`r fatalities$EVTYPE[2]`** and **`r fatalities$EVTYPE[3]`**.

### Analysis of Severe weather events economical consequences across the United States

For this analysis we're going to pay particularly attention to **property** (**PROPDMG**) and **crop** (**CROPDMG**) damages.

We need to consider the factor which need to be applied on the values (National Weather Service [Storm Data Documentation](https://d396qusza40orc.cloudfront.net/repdata%2Fpeer2_doc%2Fpd01016005curr.pdf))

To do so we'll have a look at the total numbers of  **property** and **crop** damages per event types.

```{r dmg}
prop <- aggregate(PROP ~ EVTYPE, storm, sum, na.rm = T)
crop <- aggregate(CROP ~ EVTYPE, storm, sum, na.rm = T)
# put side by side the costs
dmg <- merge(prop, crop, by = 'EVTYPE')
dmg$AMOUNT <- dmg$PROP + dmg$CROP
# sort the tables
prop <- prop[order(prop$PROP, decreasing = T), ]
crop <- crop[order(crop$CROP, decreasing = T), ]
dmg <- dmg[order(dmg$AMOUNT, decreasing = T), ]
```

```{r dmg_plot}
# reshape the table columns in a single column type
top_dmg <- reshape(head(dmg, 20), direction = 'long', varying = c('PROP', 'CROP', 'AMOUNT'), v.names = 'amount', timevar = 'type', times = c('Property', 'Crop', 'Total'))
library(ggplot2)
ggplot(data = top_dmg, aes(x = EVTYPE, y = amount, group = type, colour = type)) +
    geom_line() + geom_point() +
    labs(title = 'Weather events damages on economy', x = 'Event types', y = 'Amount in $US') +
    theme(axis.text.x=element_text(angle = 90, hjust = 1))
```

```{r dmg_tables}
head(dmg, 20)
head(prop, 10)
head(crop, 10)
```

**`r dmg$EVTYPE[1]`** are the most expensive event, followed by **`r dmg$EVTYPE[2]`** and **`r dmg$EVTYPE[3]`**.

**`r prop$EVTYPE[1]`** are the most expensive event on properties, followed by **`r prop$EVTYPE[2]`** and **`r prop$EVTYPE[3]`**.

**`r crop$EVTYPE[1]`** are the most expensive event on crop, followed by **`r crop$EVTYPE[2]`** and **`r crop$EVTYPE[3]`**.

## Results

As a conclusion we can safely consider **`r harm$EVTYPE[1]`** as the most dangerous weather event for the population.

The **`r dmg$EVTYPE[1]`** causes the greatest economic consequences on the US communities and municipalities in general and on the properties in particular. Meanwhile the **`r crop$EVTYPE[1]`** have a huge cost on the crops.
