---
title: 'Machine Learning'
author: 'Eddy ∆'
date: '22/01/2019'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# library(doRNG)
library(doParallel)
# cl <- makePSOCKcluster(detectCores() - 1)
# registerDoParallel(cl)
registerDoParallel(cores = detectCores() - 1)
getDoParWorkers()
```
# Week 2
## Quiz
### 4
```{r q4_data}
setwd('~/projects/training/datascience/machine_learning')

library(caret)
library(AppliedPredictiveModeling)
set.seed(3433)
data(AlzheimerDisease)
adData = data.frame(diagnosis,predictors)
inTrain = createDataPartition(adData$diagnosis, p = 3/4)[[1]]
training = adData[inTrain, ]
testing = adData[-inTrain, ]
```

Find all the predictor variables in the training set that begin with IL. Perform principal components on these variables with the preProcess() function from the caret package. Calculate the number of principal components needed to capture 80% of the variance. How many are there?

```{r q4_answer}
subset = training[,grep("^IL", names(training))]
preProcess(subset, thresh = 0.8, method = "pca")$numComp
```
Note : Pre-processing of Predictors transformation (centering, scaling etc.) can be estimated from the training data and applied to any data set with the same variables.

### 5
Create a training data set consisting of only the predictors with variable names beginning with IL and the diagnosis. Build two predictive models, one using the predictors as they are and one using PCA with principal components explaining 80% of the variance in the predictors. Use method="glm" in the train function. What is the accuracy of each method in the test set? Which is more accurate?

```{r q5_answer}
training_il <- training[,grep('^IL|diagnosis', x = names(training))]
model_def <- train(diagnosis ~ ., method = 'glm', data = training_il)
testing_il <- testing[,grep('^IL|diagnosis', x = names(testing))]
predic_def <- predict(model_def, newdata = testing_il)
confusionMatrix(predic_def, testing_il$diagnosis)
model_pca <- train(diagnosis ~ ., method = 'glm', preProcess = 'pca', data = training_il, trControl = trainControl(preProcOptions = list(thresh = 0.8)))
predic_pca <- predict(model_pca, newdata = testing_il)
confusionMatrix(predic_pca, testing_il$diagnosis)
```
```{r q5.1}
trainSubset = training[,grep("^IL", names(training))]
testSubset = testing[,grep("^IL", names(testing))]
pp = preProcess(trainSubset, thresh = 0.8, method = "pca")
trainTransformed <- predict(pp, trainSubset)
testTransformed <- predict(pp, testSubset)
trainSubset$diagnosis = training$diagnosis
testSubset$diagnosis = testing$diagnosis
trainTransformed$diagnosis = training$diagnosis
testTransformed$diagnosis = testing$diagnosis
glmpca = train(diagnosis ~ ., data = trainTransformed, method = "glm")
glm = train(diagnosis ~ ., data = trainSubset, method = "glm")
round(confusionMatrix(testSubset$diagnosis,predict(glm, testSubset))$overall["Accuracy"],2)
round(confusionMatrix(testTransformed$diagnosis,predict(glmpca, testTransformed))$overall["Accuracy"],2)

train_pp_pca <- preProcess(training_il[, -1], method='pca', thresh=0.8)
training_preproc_pca <- predict(train_pp_pca, training_il[, -1])
test_pp_pca <- preProcess(testing_il[, -1], method='pca', thresh=0.8)
testing_preproc_pca <- predict(test_pp_pca, testing_il[, -1])
training_preproc_pca$diagnosis <- training_il$diagnosis
testing_preproc_pca$diagnosis <- testing_il$diagnosis
model_pca2 <- train(diagnosis ~ ., data=training_preproc_pca, method='glm')
confusionMatrix(testing_preproc_pca$diagnosis, predict(model_pca2, testing_preproc_pca))
confusionMatrix(predict(model_pca2, testing_preproc_pca), testing_preproc_pca$diagnosis)
```

# Week 3
## Bagging
```{r bagging}
library(ElemStatLearn); data('ozone', package = 'ElemStatLearn')
ozone <- ozone[order(ozone$ozone), ]
head(ozone)
ll <- matrix(NA, nrow = 10, ncol = 155)
for (i in 1:10){
    ss <- sample(1:dim(ozone), replace = T)
    ozone0 <- ozone[ss, ]; ozone0 <- ozone0[order(ozone0$ozone), ]
    loess0 <- loess(temperature ~ ozone, data = ozone0, span = 0.2)
    ll[i, ] <- predict(loess0, newdata = data.frame(ozone=1:155))
}
```
```{r}
plot(ozone$ozone, ozone$temperature, pch = 19, cex = 0.5)
for (i in 1:10) {lines(1:155, ll[i, ], col = 'grey', lwd = 2)}
lines(1:155, apply(ll, 2, mean), col = 'red', lwd = 2)
```

## Random Forest
```{r random_forest}
library(ggplot2)
library(caret)
library(randomForest)
data(iris)
inTrain <- createDataPartition(y = iris$Species, p = 0.7, list = F)
training <- iris[inTrain, ]; testing <- iris[-inTrain, ]
modFit <- train(Species ~ ., data = training, method = 'rf', prox = T)
modFit
getTree(modFit$finalModel, k = 2)
irisP <- classCenter(training[, c(3, 4)], training$Species, modFit$finalModel$prox)
irisP <- as.data.frame(irisP); irisP$Species <- rownames(irisP)
p <- qplot(Petal.Width, Petal.Length, col = Species, data = training)
p + geom_point(aes(x = Petal.Width, y = Petal.Length, col = Species), size = 5, shape = 4, data = irisP)
pred <- predict(modFit, testing); testing$predRight <- pred == testing$Species
table(pred, testing$Species)
qplot(Petal.Width, Petal.Length, colour = predRight, data = testing, main = 'newdata predictions')
```
## Boosting
```{r boosting, message = FALSE}
library(ISLR); library(ggplot2); library(caret); data(Wage)
Wage <- subset(Wage, select = -c(logwage)) # NOTE : subset remove logwage column
inTrain <- createDataPartition(y = Wage$wage, p = 0.7, list = F)
training <- Wage[inTrain, ]; testing <- Wage[-inTrain, ]
modFit <- train(wage ~ ., data = training, method = 'gbm', verbose = F) # NOTE : gbm boost with trees 
```
```{r print_it, echo=FALSE}
print(modFit)
```

## Model based prediction
```{r discriminant}
names(iris)
table(iris$Species)
inTrain <- createDataPartition(y = iris$Species, p = 0.7, list = F)
training <- iris[inTrain, ]; testing <- iris[-inTrain, ]
dim(training); dim(testing)
modlda = train(Species ~ ., data = training, method = 'lda')
modnb = train(Species ~ ., data = training, method = 'nb')
plda = predict(modlda, testing); pnb = predict(modnb, testing)
table(plda, pnb)
equalPredictions = (plda == pnb)
qplot(Petal.Width, Sepal.Width, colour = equalPredictions, data = testing)
```

## Quiz
### 1
1. Subset the data to a training set and testing set based on the Case variable in the data set.

2. Set the seed to 125 and fit a CART model with the rpart method using all predictor variables and default caret settings.

3. In the final model what would be the final model prediction for cases with the following variable values:

a. TotalIntench2 = 23,000; FiberWidthCh1 = 10; PerimStatusCh1=2

b. TotalIntench2 = 50,000; FiberWidthCh1 = 10;VarIntenCh4 = 100

c. TotalIntench2 = 57,000; FiberWidthCh1 = 8;VarIntenCh4 = 100

d. FiberWidthCh1 = 8;VarIntenCh4 = 100; PerimStatusCh1=2
```{r quiz3}
library(rattle)
library(AppliedPredictiveModeling)
data(segmentationOriginal)
set.seed(125)
## inTrain <- createDataPartition(y = segmentationOriginal$Case, list = F)
training <- segmentationOriginal[segmentationOriginal$Case == 'Train', ]
testing <- segmentationOriginal[segmentationOriginal$Case == 'Test', ]
modFit <- train(Class ~ ., data = training, method = 'rpart')
modFit$finalModel
fancyRpartPlot(modFit$finalModel)
# pcart <- predict(modFit, testing)
```
### 2
If K is small in a K-fold cross validation is the bias in the estimate of out-of-sample (test set) accuracy smaller or bigger? If K is small is the variance in the estimate of out-of-sample (test set) accuracy smaller or bigger. Is K large or small in leave one out cross validation?

The bias is smaller and the variance is bigger. Under leave one out cross validation K is equal to one.
### 3
These data contain information on 572 different Italian olive oils from multiple regions in Italy. Fit a classification tree where Area is the outcome variable. Then predict the value of area for the following data frame using the tree command with all defaults.

What is the resulting prediction? Is the resulting prediction strange? Why or why not?
```{r}
library(pgmm)
data(olive)
stargazer::stargazer(olive[1:10, ], title = 'Olive trees data set') # NOTE : , type = 'html'
olive = olive[,-1]
# fancyRpartPlot(rpart(Area ~ ., data = olive)$finalModel)
# fancyRpartPlot(train(Area ~ ., data = olive, method = 'rpart')$finalModel)
newdata = as.data.frame(t(colMeans(olive)))
inTrain <- createDataPartition(y = olive$Area, p = 0.7, list = F)
training <- olive[inTrain, ]; testing <- olive[-inTrain, ]
modFit <- train(Area ~ ., data = training, method = 'rpart')
modFit$finalModel
fancyRpartPlot(modFit$finalModel)
olive2 <- transform(olive, Area = factor(Area, 1:9, c('North Apulia', 'Calabria', 'South Apulia', 'Sicily', 'Inland Sardinia', 'Costal Sardinia', 'East Liguria', 'West Liguria', 'Umbria')))
inTrain <- createDataPartition(y = olive2$Area, p = 0.7, list = F)
training <- olive2[inTrain, ]; testing <- olive2[-inTrain, ]
modFit <- train(Area ~ ., data = training, method = 'rpart')
modFit$finalModel
fancyRpartPlot(modFit$finalModel)
```
2.783. It is strange because Area should be a qualitative variable - but tree is reporting the average value of Area as a numeric variable in the leaf predicted for newdata

### 4
Load the South Africa Heart Disease Data and create training and test sets with the following code:
```{r}
library(ElemStatLearn)
data(SAheart)
set.seed(8484)
train = sample(1:dim(SAheart)[1],size=dim(SAheart)[1]/2,replace=F)
trainSA = SAheart[train,]
testSA = SAheart[-train,]
```
Then set the seed to 13234 and fit a logistic regression model (method="glm", be sure to specify family="binomial") with Coronary Heart Disease (chd) as the outcome and age at onset, current alcohol consumption, obesity levels, cumulative tabacco, type-A behavior, and low density lipoprotein cholesterol as predictors. Calculate the misclassification rate for your model using this function and a prediction on the "response" scale:
```{r}
missClass = function(values,prediction){sum(((prediction > 0.5)*1) != values)/length(values)}
```
What is the misclassification rate on the training set? What is the misclassification rate on the test set?
```{r}
set.seed(13234)
modelFit <- train(chd ~ age + alcohol + obesity + tobacco + typea + ldl, data = trainSA, method = 'glm', family = 'binomial')
predictTest <- predict(modelFit, testSA)
missClass(testSA$chd, predictTest)
predictTrain <- predict(modelFit)
missClass(trainSA$chd, predictTrain)
```
Test : `r missClass(testSA$chd, predictTest)`, Train : `r missClass(trainSA$chd, predictTrain)`.

### 5
Load the vowel.train and vowel.test data sets:
```{r}
library(ElemStatLearn)
data(vowel.train)
data(vowel.test)
training <- transform(vowel.train, y = factor(y))
testing <- transform(vowel.test, y = factor(y))
```
Calculate the variable importance using the varImp function in the caret package. What is the order of variable importance?
```{r}
set.seed(33833)
modFit <- train(y ~ ., data = training, method = 'rf', prox = T)
# modFit <- randomForest(y ~ ., data = training)
# modFit
predictTest <- predict(modFit, testing)
# getTree(modFit$finalModel, k = 2)
varImp(modFit)
```

# Week 4
## Combining predictors
```{r combine_predictors}
library(ISLR); data(Wage); library(ggplot2); library(caret)
Wage <- subset(Wage, select = -c(logwage))
# create a building data set and validation set
inBuild <- createDataPartition(y = Wage$wage, p = 0.7, list = F)
validation <- Wage[-inBuild, ]; buildData <- Wage[inBuild, ]
inTrain <- createDataPartition(y = buildData$wage, p = 0.7, list = F)
training <- buildData[inTrain, ]; testing <- buildData[-inTrain, ]
dim(training)
dim(testing)
dim(validation)
mod1 <- train(wage ~ ., data = training, method = 'glm')
# trainControl cv (crossvalidation)
mod2 <- train(wage ~ ., data = training, method = 'rf', trControl = trainControl(method = 'cv'), number = 3)
pred1 <- predict(mod1, testing); pred2 <- predict(mod2, testing)
qplot(pred1, pred2, colour = wage, data = testing)
predDF <- data.frame(pred1, pred2, wage = testing$wage)
combModFit <- train(wage ~ ., data = predDF, method = 'gam')
combPred <- predict(combModFit, predDF)
sqrt(sum((pred1 - testing$wage)^2))
sqrt(sum((pred2 - testing$wage)^2))
sqrt(sum((combPred - testing$wage)^2))
pred1V <- predict(mod1, validation); pred2V <- predict(mod2, validation)
predVDF <- data.frame(pred1 = pred1V, pred2 = pred2V)
combPredV <- predict(combModFit, predVDF)
sqrt(sum((pred1V - validation$wage)^2))
sqrt(sum((pred2V - validation$wage)^2))
sqrt(sum((combPredV - validation$wage)^2))
```

## Forecasting
```{r forecasting}
library(quantmod)
from.dat <- as.Date('01/01/08', format = '%m/%d/%y')
to.dat <- as.Date('12/31/13', format = '%m/%d/%y')
# getSymbols('GOOG', src = 'google', from = from.dat, to = to.dat)
# rm(GOOG)
getSymbols('GOOG', src = 'yahoo', from = from.dat, to = to.dat)
head(GOOG)
# convert to a monthly time series
mGoog <- to.monthly(GOOG)
# Take the opening information
googOpen <- Op(mGoog)
# create a time series object
ts1 <- ts(googOpen, frequency = 12)
plot(ts1, xlab = 'Year+1', ylab = 'GOOG')
# Decomposition of additive time series
# last random kind of cyclical pattern
plot(decompose(ts1), xlab = 'Year+1')
ts1Train <- window(ts1, start = 1, end = 5)
ts1Test <- window(ts1, start = 5, end = (7 - 0.01))
ts1Train
plot(ts1Train)
library(forecast)
lines(ma(ts1Train, order = 3), col = 'red')
ets1 <- ets(ts1Train, model = 'MMM')
fcast <- forecast(ets1)
plot(fcast); lines(ts1Test, col = 'red')
accuracy(fcast, ts1Test)
```
## Unsupervised prediction
```{r unsupervised}
data(iris)
inTrain <- createDataPartition(y = iris$Species, p = 0.7, list = F)
training <- iris[inTrain, ]; testing <- iris[-inTrain, ]
dim(training); dim(testing)
KMeans1 <- kmeans(subset(training, select = -c(Species)), centers = 3)
training$clusters <- as.factor(KMeans1$cluster)
qplot(Petal.Width, Petal.Length, colour = clusters, data = training)
table(KMeans1$cluster, training$Species)
modFit <- train(clusters ~ ., data = subset(training, select = -c(Species)), method = 'rpart')
table(predict(modFit, training), training$Species)
testClusterPred <- predict(modFit, testing)
table(testClusterPred, testing$Species)
```
## Quiz
### 1 combine 2 models
```{r}
library(ElemStatLearn)
data(vowel.train)
data(vowel.test)
```
Set the variable y to be a factor variable in both the training and test set. Then set the seed to 33833. Fit (1) a random forest predictor relating the factor variable y to the remaining variables and (2) a boosted predictor using the "gbm" method. Fit these both with the train() command in the caret package.

What are the accuracies for the two approaches on the test data set? What is the accuracy among the test set samples where the two methods agree?
```{r}
training <- transform(vowel.train, y = factor(y))
testing <- transform(vowel.test, y = factor(y))
set.seed(33833)
modRf <- train(y ~ ., data = training, method = 'rf') # NOTE : random forest, prox = T
modGbm <- train(y ~ ., data = training, method = 'gbm', verbose = F) # NOTE : gbm boost with trees
predRf <- predict(modRf, testing)
predGbm <- predict(modGbm, testing)
confusionMatrix(testing$y, predRf)$overall['Accuracy']
confusionMatrix(testing$y, predGbm)$overall['Accuracy']
predMatch <- predRf == predGbm
confusionMatrix(testing$y[predMatch], predRf[predMatch])$overall['Accuracy']
# qplot(pred1, pred2, colour = y, data = testing)
# varImp(modFit)
```
### 2 combine 3 models
Load the Alzheimer's data using the following commands
```{r}
library(gbm)
set.seed(3433)
library(AppliedPredictiveModeling)
data(AlzheimerDisease)
adData = data.frame(diagnosis,predictors)
inTrain = createDataPartition(adData$diagnosis, p = 3/4)[[1]]
training = adData[inTrain, ]
testing = adData[-inTrain, ]
```
Set the seed to 62433 and predict diagnosis with all the other variables using a random forest ("rf"), boosted trees ("gbm") and linear discriminant analysis ("lda") model. Stack the predictions together using random forests ("rf"). What is the resulting accuracy on the test set? Is it better or worse than each of the individual predictions?
```{r}
set.seed(62433)
modRf <- train(diagnosis ~ ., data = training, method = 'rf') # NOTE : random forest, prox = T
modGbm <- train(diagnosis ~ ., data = training, method = 'gbm', verbose = F) # NOTE : gbm boost with trees
modLda = train(diagnosis ~ ., data = training, method = 'lda')
predRf <- predict(modRf, testing)
predGbm <- predict(modGbm, testing)
predLda <- predict(modLda, testing)
# merging the pred together
predDF <- data.frame(predRf, predGbm, predLda, diagnosis = testing$diagnosis)
combModFit <- train(diagnosis ~ ., data = predDF, method = 'rf')
# combPred <- predict(combModFit, predDF)
combPred <- predict(combModFit, testing)
confusionMatrix(testing$diagnosis, predRf)$overall['Accuracy']
confusionMatrix(testing$diagnosis, predGbm)$overall['Accuracy']
confusionMatrix(testing$diagnosis, predLda)$overall['Accuracy']
confusionMatrix(testing$diagnosis, combPred)$overall['Accuracy']
# sqrt(sum((pred1 - testing$y)^2))
# sqrt(sum((pred2 - testing$y)^2))
# sqrt(sum((combPred - testing$y)^2))
```
### 3 lasso
Load the concrete data with the commands:
```{r}
set.seed(3523)
library(AppliedPredictiveModeling)
data(concrete)
inTrain = createDataPartition(concrete$CompressiveStrength, p = 3/4)[[1]]
training = concrete[inTrain, ]
testing = concrete[-inTrain, ]
```
Set the seed to 233 and fit a lasso model to predict Compressive Strength. Which variable is the last coefficient to be set to zero as the penalty increases? (Hint: it may be useful to look up ?plot.enet).
```{r}
set.seed(233)
modLasso <- train(CompressiveStrength ~ ., data = training, method = 'lasso')
plot.enet(modLasso$finalModel, xvar = 'penalty', use.color = TRUE)
# Cement
```
### 4 forecast
```{r load_file}
get_file <- function(file_url, data_dir = 'data'){
    if(!dir.exists(data_dir)) dir.create(data_dir)
    # file_name <- basename(file_url)
    # file_path <- file.path(data_dir, file_name)
    file_path <- file.path(data_dir, basename(file_url))
    if(!file.exists(file_path)) download.file(file_url, file_path, quiet = TRUE)
    # if(!file.exists(file_path))
    #     stop('Unable to download file ', file_url)
    # return the archive file path
    file_path
}
```

Load the data on the number of visitors to the instructors blog from here:

[https://d396qusza40orc.cloudfront.net/predmachlearn/gaData.csv](https://d396qusza40orc.cloudfront.net/predmachlearn/gaData.csv)

Using the commands:
```{r}
library(lubridate) # For year() function below
# dat = read.csv('~/Desktop/gaData.csv')
dat = read.csv(get_file('https://d396qusza40orc.cloudfront.net/predmachlearn/gaData.csv'))
training = dat[year(dat$date) < 2012, ]
testing = dat[year(dat$date) > 2011, ]
tstrain = ts(training$visitsTumblr)
```
Fit a model using the bats() function in the forecast package to the training time series. Then forecast this model for the remaining time points. For how many of the testing points is the true value within the 95% prediction interval bounds?
```{r}
modFit <- bats(training$visitsTumblr)
forecast95 <- forecast(modFit, level = 95, h = nrow(testing))
names(forecast95)
in95 <- sum(testing$visitsTumblr > forecast95$lower & testing$visitsTumblr < forecast95$upper)
in95 / nrow(testing)
```
### 5
Load the concrete data with the commands:
```{r}
set.seed(3523)
library(AppliedPredictiveModeling)
data(concrete)
inTrain = createDataPartition(concrete$CompressiveStrength, p = 3/4)[[1]]
training = concrete[inTrain, ]
testing = concrete[-inTrain, ]
```
Set the seed to 325 and fit a support vector machine using the e1071 package to predict Compressive Strength using the default settings. Predict on the testing set. What is the RMSE?
```{r}
set.seed(325)
library(e1071)
modSvm <- svm(CompressiveStrength ~ ., data = training)
predSvm <- predict(modSvm, testing)
accSvm <- accuracy(predSvm, testing$CompressiveStrength)
data.frame(accSvm)['RMSE']
```

# WLE
[Caret Package – A Practical Guide to Machine Learning in R](https://www.machinelearningplus.com/machine-learning/caret-package)

```{r}
set.seed(99)
# NOTE : same `randomForest` function <- modRf <- randomForest(classe ~ ., data = trainingRf)
# table(pred, transform(validationRf, predRight = predRf == validationRf$classe)$classe)

# modGlm <- train(classe ~ ., data = trainingRf, method = 'glm', preProcess = 'pca', trControl = trainControl(preProcOptions = list(thresh = 0.8)))

# prep <- preProcess(trainingRf[, -53], thresh = 0.8, method = 'pca')
# training_prep <- predict(prep, trainingRf)
# modGlm <- train(classe ~ ., data = training_prep, method = 'glm')

## Random forest
library(randomForest)
benchmark(modRfTrain <- train(classe ~ ., data = trainingRf, method = 'rf', prox = T))
benchmark(modRfRF <- randomForest(classe ~ ., data = trainingRf))
benchmark(modRfCv <- train(classe ~ ., data = trainingRf, method = 'rf', prox = T, trControl = trainControl(method = 'cv', number = 5, allowParallel = TRUE)))
benchmark(modRfBoot <- train(classe ~ ., data = trainingRf, method = 'rf', prox = T, trControl = trainControl(method = 'boot', number = 1, allowParallel = TRUE)))

## Decision tree
library(rpart)
benchmark(modRpartTrain <- train(classe ~ ., data = training, method = 'rpart'))
benchmark(modRpartRP <- rpart(classe ~ ., data = training, method = 'class'))
benchmark(modRpartTrain2 <- train(classe ~ ., method = 'rpart', data = trainingRpart, trControl = trainControl(method = 'none'), tuneGrid = data.frame(cp = 0.01)))
```
```{r training_day}
set.seed(9999)
b <- date()
modRfx <- train(classe ~ ., data = training, method = 'rf', prox = T, trControl = trainControl(method = 'cv', number = 10, allowParallel = TRUE))
e <- date()
b;e
beep()
modRfx
modRfx$resample
confusionMatrix.train(modRfx)
confusionMatrix(predict(modRfx, trainingRpart), trainingRpart$classe)
```

```{r}
set.seed(9999)
b <- date()
# modRpartx <- train(classe ~ ., method = 'rpart', data = training, trControl = trainControl(method = 'none', allowParallel = TRUE, verboseIter = FALSE), tuneGrid = data.frame(cp = 0.0001))
modRpartx <- train(classe ~ ., method = 'rpart', data = training, trControl = trainControl(method = 'cv', number = 10, allowParallel = TRUE, verboseIter = FALSE), tuneGrid = data.frame(cp = 0.0001))
e <- date()
beep()
modRpartx
confusionMatrix.train(modRpartx)
confusionMatrix(predict(modRpartx, validation), validation$classe)
```

## Gradient Boosting Model
```{r gbm, cache = TRUE}
modGbm <- train(classe ~ ., method = 'gbm', data = training, trControl = trainControl(method = 'cv', number = 10, allowParallel = TRUE, verboseIter = FALSE), verbose = FALSE)
modGbm
confusionMatrix.train(modGbm)
cmGbm <- confusionMatrix(predict(modGbm, validation), validation$classe)
cmGbm
```

## Multinomial Logistic Regression
```{r gbm, cache = TRUE}
modMlr <- multinom(classe ~ ., data = training)
modMlr
cmMlr <- confusionMatrix(predict(modMlr, validation), validation$classe)
cmMlr
```

```{r echo=FALSE}
# stopCluster(cl)
stopImplicitCluster()
```

```{r}
# library(DBI)
# srv <- dbConnect(RSQLite::SQLite(), dbname = "chinook.db")
# query <- "SELECT * FROM tracks"
```
```{r}
library(RPostgreSQL)
library(DBI)
pgdrv <- dbDriver(drvName = 'PostgreSQL')
db <-DBI::dbConnect(pgdrv, dbname = 'postgres', host = 'localhost', port = 5432)
# db <-DBI::dbConnect(pgdrv, dbname = 'postgres', host = 'localhost', port = 5432, user = 'postgres', password = getPass('Enter Password:'), )
# DBI::dbWriteTable(db, "mtcars", mtcars)
query <- 'select 1'
# DBI::dbGetQuery(db, 'SELECT * FROM MTCARS')
DBI::dbGetQuery(db, 'select 1')

```
```{sql, connection = db}
select 1
```
```{sql, connection = db}
EXEC (?query)
```
```{r sql_out, connection = db}
DBI::dbDisconnect(db)
```

```{r}
# install.packages("MonetDBLite")

library(DBI)
dbdir <- tempdir()
con <- dbConnect(MonetDBLite::MonetDBLite())
dbWriteTable(con, "mtcars", mtcars)
dbGetQuery(con, "SELECT MAX(mpg) FROM mtcars WHERE cyl = 8")

library(dplyr)
ms <- MonetDBLite::src_monetdblite(dbdir)
mt <- tbl(ms, "mtcars")
mt %>% filter(cyl == 8) %>% summarise(max(mpg))
# shut down the database
dbDisconnect(con, shutdown = TRUE)
MonetDBLite::monetdblite_shutdown()

# To connect to a database first create a src:
dbdir <- file.path(tempdir(), "dplyrdir")
my_db <- MonetDBLite::src_monetdblite(dbdir)
# copy some data to DB
my_mtcars  <- copy_to(my_db, mtcars)

# create table object
my_mtcars2 <- tbl(my_db, 'mtcars')
# shut down the database
dbDisconnect(con, shutdown = TRUE)
MonetDBLite::monetdblite_shutdown()
```
