---
title: "College Major"
author: "Eddy ∆"
date: "20/11/2018"
output:
  pdf_document: default
  html_document: default
subtitle: How income varies across college major categories
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load data

```{r load_data}
install.packages("devtools")
install.packages("broom")
devtools::install_github("jhudsl/collegeIncome")
library(collegeIncome)
data(college)
```

A codebook for the dataset is given below:

* rank: Rank by median earnings
* major_code: Major code
* major: Major description
* major_category: Category of major
* total: Total number of people with major
* sample_size: Sample size of full-time, year-round individuals used for income/earnings estimates: p25th, median, p75th
* p25th: 25th percentile of earnings
* median: Median earnings of full-time, year-round workers
* p75th: 75th percentile of earnings
* perc_men: % men with major (out of total)
* perc_women: % women with major (out of total)
* perc_employed: % employed (out of total)
* perc_employed_fulltime: % employed 35 hours or more (out of employed)
* perc_employed_parttime: % employed less than 35 hours (out of employed)
* perc_employed_fulltime_yearround: % employed at least 50 weeks and at least 35 hours (out of employed and full-time)
* perc_unemployed: % unemployed (out of employed)
* perc_college_jobs: % with job requiring a college degree (out of employed)
* perc_non_college_jobs: % with job not requiring a college degree (out of employed)
* perc_low_wage_jobs: % in low-wage service jobs (out of total)


# Exploratory data analysis

```{r college}
str(college)
summary(college)
```

Tidy the data by changing `major_code`, `major`, and `major_category` to factor
```{r college_major}
college_major <- transform(
    college,
    major_code = factor(major_code),
    major = factor(major),
    major_category = factor(major_category)
)
```

## Data analysis

You can also embed plots, for example:

```{r ggpairs, echo=FALSE}
ggpairs(college_major[, c('major_category', 'median')], lower = list(continuous = wrap('smooth', method = 'loess')), cardinality_threshold = 16)
```

```{r cfit}
cfit = lm(median ~ major_category, college_major)
summary(cfit)
anova(cfit)
broom::tidy(cfit)
```

## Based on your analysis, would you conclude that there is a significant association between college major category and income?
No

## Please type a few sentences describing your results.

A simple linear regression between major_category (Category of major) and median (Median earnings of full-time, year-round workers) came out with a high p-value for the t test score. This outcome is also consitent with an anova analysis where the p-value is also high for the F test score. Those results makes me infer that there are no significant relationship between the median earning and the major's category.
Hence deeper analysis would interesting to be done to check which other variable would be susceptible to be more correlated to the major's category.

Out of curiosity let's explore other variables

```{r pairs, echo=FALSE}
pairs(college_major[, c('median', 'perc_men', 'perc_employed', 'perc_employed_parttime', 'perc_low_wage_jobs')],panel=panel.smooth, pch=16, cex=0.5, gap=0.25, lwd=2, las=1, cex.axis=0.7)
```

