---
title: 'Basic Inferential Data Analysis'
subtitle: 'The Effect of Vitamin C on Tooth Growth in Guinea Pigs'
author: "Eddy Delta"
date: "01/09/2018"
output: pdf_document
geometry: margin=1.2cm
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Basic summary of the data

The ToothGrowth data in the R datasets package contains the length of odontoblasts (cells responsible for tooth growth) in 60 guinea pigs. Each animal received one of three dose levels of vitamin C (0.5, 1, and 2 mg/day) by one of two delivery methods, orange juice or ascorbic acid (a form of vitamin C and coded as VC).

# Overview

```{r libraries}
library(ggplot2)
library(datasets)
data(ToothGrowth)
```

## Data distribution

```{r distribution}
ggplot(ToothGrowth, aes(x = factor(dose), y = len, fill = factor(dose))) +
  geom_boxplot() + 
  facet_grid(.~supp) +
  xlab("Dosage") +   
  ylab("Teeth Growth") +  
  ggtitle("Guinea pigs teeth Growth by supplement type")
```

Quick look at the data contents

```{r content}
str(ToothGrowth)
```

Some meaning full values about the data content

```{r overview}
summary(ToothGrowth)
```

The data shows a possible correlation between the teeth growth and the dosage on both supplement type.

# Analysis

## Dosage influence

To analyse the correlation between the teeth growth and the dosage, we'll consider the differents pairs of dosage. With $H_{0} : corr(dosage_{x}, dosage_{y}) = 0$ and  $H_{A} : corr(dosage_{x}, dosage_{y}) \ne 0$

### 0.5 vs 1.0

```{r dosage_1}
t.test(len~dose, data=subset(ToothGrowth, dose %in% c(0.5, 1.0)), alternative = 'two.sided', paired=FALSE)
```

### 0.5 vs 2.0

```{r dosage_2}
t.test(len~dose, data=subset(ToothGrowth, dose %in% c(0.5, 2.0)), alternative = 'two.sided', paired=FALSE)
```

### 1.0 vs 2.0

```{r dosage_3}
t.test(len~dose, data=subset(ToothGrowth, dose %in% c(1.0, 2.0)), alternative = 'two.sided', paired=FALSE)
```

0 is out of the 95% confidence interval in every cases so $H_{0}$ is rejected, which makes us conclude with 95% of confidence that there is a correlation between teeth growth and Vitamin C dose.

## Delivery methods influence

Let's analyse the correlation between the teeth growth and the delivery methods. With $H_{0} : corr(delivery_{OJ}, delivery_{VC}) = 0$ and  $H_{A} : corr(delivery_{OJ}, delivery_{VC}) \ne 0$

```{r delivery}
t.test(len ~ supp, paired = F, var.equal = F, data = ToothGrowth)
```

The $H_{0}$ can't be rejected with the 95% of confidense because the 95% confidence interval contains 0. We then can't conclude with 95% of confidence that there is a correlation between the teeth growth and the delivery method.

# Conclusion

We can assume assume with 95% of confidence the the teeth growth is positively correlated to the teeth growth. However there is no enough data to conclude any correlation between the teeth growth and the delivery method with 95% confidence.
